### 2025-03-15 paper/preprint/paper.tex
done

\begin{itemize}
\tightlist
\item
  \(y^+\): The target class and also the index of the target class.
\item
  \(y^-\): The non-target class and also the index of non-the target
  class.
\item
  \(\mathbf{y}^+\): The one-hot encoded output vector for the target
  class.
\item
  \(\theta\): Model parameters (unspecified).
\item
  \(\Theta\): Matrix of parameters.
\end{itemize}

@@

\begin{equation}\phantomsection\label{eq-mmd}{
\begin{aligned}
MMD({X}^\prime,\tilde{X}^\prime) &= \frac{1}{m(m-1)}\sum_{i=1}^m\sum_{j\neq i}^m k(x_i,x_j) \\ &+ \frac{1}{n(n-1)}\sum_{i=1}^n\sum_{j\neq i}^n k(\tilde{x}_i,\tilde{x}_j) \\ &- \frac{2}{mn}\sum_{i=1}^m\sum_{j=1}^n k(x_i,\tilde{x}_j)
\end{aligned}
}\end{equation}

In our implementation, Equation~\ref{eq-mmd} is by default applied to
the entire subset of the training data for which \(y=y^+\).

@@

\subsection{Training Details}\label{sec-app-training}

In this section, we describe the training procedure in detail. While the
details laid out here are not crucial for understanding our proposed
approach, they are of importance to anyone looking to implement
counterfactual training.

\subsubsection{Default Hyperparameters}\label{default-hyperparameters}

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-train-default}: Training Phase}]

@@

Note~\ref{nte-eval-default} presents the default hyperparameters used
during evaluation.

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-eval-default}: Evaluation Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-gen-params-final-run-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-gen-params-final-run-eval}: Evaluation Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-pen-final-run-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-pen-final-run-eval}: Evaluation Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-train-final-run-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-train-final-run-eval}: Evaluation Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-warning-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-warning-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Package Version (Reproducibility)}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-tune-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-tune-eval}: Evaluation Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-tune_lr-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, breakable, titlerule=0mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, toptitle=1mm, bottomtitle=1mm, title={Note \ref*{nte-tune_lr-eval}: Evaluation Phase}]

>>>>

Below we provide an overview of some notation used frequently throughout
the paper:

\begin{itemize}
\tightlist
\item
  \(y^+\): The target class and also the index of the target class.
\item
  \(y^-\): The non-target class and also the index of non-the target
  class.
\item
  \(\mathbf{x}\): a single training sample.
\item
  \(\mathbf{x}^\prime\): a counterfactual.
\item
  \(\mathbf{x}^+\): a training sample in the target class
  (ground-truth).
\item
  \(\mathbf{y}^+\): The one-hot encoded output vector for the target
  class.
\item
  \(\theta\): Model parameters (unspecified).
\item
  \(\Theta\): Matrix of parameters.
\item
  \(\mathbf{M}(\cdot)\): linear predictions (logits) of the classifier.
\end{itemize}

@@

Maximum mean discrepancy is defined as follows,

\begin{equation}\phantomsection\label{eq-mmd}{
\begin{aligned}
\text{MMD}({X}^\prime,\tilde{X}^\prime) &= \frac{1}{m(m-1)}\sum_{i=1}^m\sum_{j\neq i}^m k(x_i,x_j) \\ &+ \frac{1}{n(n-1)}\sum_{i=1}^n\sum_{j\neq i}^n k(\tilde{x}_i,\tilde{x}_j) \\ &- \frac{2}{mn}\sum_{i=1}^m\sum_{j=1}^n k(x_i,\tilde{x}_j)
\end{aligned}
}\end{equation}

where \(k(\cdot,\cdot)\) is a kernel function
(\citeproc{ref-gretton2012kernel}{Gretton et al. 2012}). We make use of
a Gaussian kernel with a constant length-scale parameter of \(0.5\). In
our implementation, Equation~\ref{eq-mmd} is by default applied to the
entire subset of the training data for which \(y=y^+\).

@@

\subsection{Training Hyperparameters}\label{sec-app-training}

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-train-default}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

For all of our evaluations, we proceed as follows: for each experiment
setting we generate multiple counterfactuals (``No.~Counterfactuals''),
randomly choosing the factual and target class each time
(Note~\ref{nte-eval-default}). We do this across multiple rounds
(``No.~Runs'') with different random seeds to account for stochasticity
(Note~\ref{nte-eval-default}). This is in line with standard practice in
the related literature on CE. Note~\ref{nte-eval-default} presents the
default hyperparameters used during evaluation. For our final results
presented in the main paper, we rely on held out test sets to sample
factuals (and outputs for our performance metrics). For tuning purposes
we rely on training or validation sets.

\subsubsection{Robust Accuracy}\label{robust-accuracy}

To evaluate robust accuracy (Acc.\(^*\)), we use the Fast Gradient Sign
Method (FGSM) to perturb test samples
(\citeproc{ref-goodfellow2014explaining}{Goodfellow, Shlens, and Szegedy
2015}). For the main results, we have set the perturbation size to
\(\epsilon=0.03\). We have also tested other perturbation sizes, as well
as randomly perturbed data. Although not reported here, we have
consistently found strong outperformance of CT compared to the weak
baseline.

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-eval-default}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-gen-params-final-run-train}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-gen-params-final-run-eval}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-pen-final-run-train}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-pen-final-run-eval}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-train-final-run-train}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-train-final-run-eval}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-warning-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-warning-color-frame, title={Package Version (Reproducibility)}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-tune-train}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-tune-eval}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-tune_lr-train}: Training Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, colback=white, titlerule=0mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, title={Note \ref*{nte-tune_lr-eval}: Evaluation Phase}, bottomtitle=1mm, arc=.35mm, left=2mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, opacityback=0, breakable]

### 2025-03-15 paper/sections/other/notation.qmd
done with grid search appendix
done

- $y^+$: The target class and also the index of the target class.
- $y^-$: The non-target class and also the index of non-the target class.
- $\mathbf{y}^+$: The one-hot encoded output vector for the target class. 
- $\theta$: Model parameters (unspecified).
- $\Theta$: Matrix of parameters.

>>>>

Below we provide an overview of some notation used frequently throughout the paper:

- $y^+$: The target class and also the index of the target class.
- $y^-$: The non-target class and also the index of non-the target class.
- $\mathbf{x}$: a single training sample.
- $\mathbf{x}^\prime$: a counterfactual. 
- $\mathbf{x}^+$: a training sample in the target class (ground-truth).
- $\mathbf{y}^+$: The one-hot encoded output vector for the target class. 
- $\theta$: Model parameters (unspecified).
- $\Theta$: Matrix of parameters. 
- $\mathbf{M}(\cdot)$: linear predictions (logits) of the classifier.

## Other Technical Details

Maximum mean discrepancy is defined as follows,

$$
\begin{aligned}
\text{MMD}({X}^\prime,\tilde{X}^\prime) &= \frac{1}{m(m-1)}\sum_{i=1}^m\sum_{j\neq i}^m k(x_i,x_j) \\ &+ \frac{1}{n(n-1)}\sum_{i=1}^n\sum_{j\neq i}^n k(\tilde{x}_i,\tilde{x}_j) \\ &- \frac{2}{mn}\sum_{i=1}^m\sum_{j=1}^n k(x_i,\tilde{x}_j)
\end{aligned}
$$ {#eq-mmd}

where $k(\cdot,\cdot)$ is a kernel function [@gretton2012kernel]. We make use of a Gaussian kernel with a constant length-scale parameter of $0.5$. In our implementation, @eq-mmd is by default applied to the entire subset of the training data for which $y=y^+$.
